### 2015-07-06 narrative/data-reduction.tex
Another update

Ideally, the scientist would perform the entire analysis stage {\em in situ} to
extract the relevant information and, in the unlikely event of this being
possible, would effectively circumvent the large data issue completely.  The
problem of course is that this is unlikely to be possible since, by their
nature, large scale simulations aim to discover new information that is often
hidden in the form of higher order effects amongst the data deluge. In
particular, this means if data thinning or truncation is applied willy nilly
then one is in danger of effectively throwing out the baby with the bath water.
One cannot store the entire data set in easily accessible source, due to sheer
size, yet one cannot reduce it prior to archival without risking losing the
information one is trying to extract.

Viewed in this way, the problem would appear intractable. However, the above
discussion fails to recognize that much of the data is redundant in the
information theoretic sense.  That is to say, the amount of information
contained in the data is often significantly less than the amount of data.  The
difficulty stems from the fact that one does not know about this redundancy in
advance.

Once we realize that data can be reduced from many different techniques, we
understand that with more application knowledge, we can reduce the total data
size much more than without knowledge, so we need to be able to design the SSIO
layers to be able to have generic data reduction techniques, and generic data
re-organization techniques. But we need to system to allow give us more
semantic knowledge of the data so that we can either: 1) Use an approach which
can audit the calculation and save values of the objects when the auditor is
off more than a given amount, and we can also save the auditor in a docker
container, which can then be executed as part of SSIO software layer.

Recognizing access patterns in an application is a key to potentially reducing
future file read time. Scientific applications often read and analyze data
alternately, thus by overlapping the two phases with prefetching can
significantly reduce the overall execution time of the application. Accurate
prefetching can be achieved with access pattern analysis. This requires an SSIO
layer which can prefetch data in a asynchronous fashion to take the bytes on
the file.

>>>>

Ideally, the scientist would perform the entire analysis stage {\em in situ} to
extract the relevant information and, in the unlikely event of this being
possible, would effectively circumvent the large data issue completely.  The
problem of course is that this is unlikely to be possible since, by their
nature, large scale simulations aim to discover new information that is often
hidden in the form of higher order effects amongst the data deluge. In
particular, this means if data thinning or truncation is applied willy nilly
then one is in danger of effectively throwing out the baby with the bath water.
One cannot store the entire data set in an easily accessible source, due to
sheer size, yet one cannot reduce it prior to archival without risking losing
the information one is trying to extract.  Viewed in this way, the problem
would appear intractable. However, the above discussion fails to recognize that
much of the data is redundant in the information theoretic sense.  That is to
say, the amount of information contained in the data is often significantly
less than the amount of data.  The difficulty stems from the fact that one does
not know about this redundancy in advance \emph{without the benefit of 
priori knowledge}. This knowledge can often be provided by the user or can 
be acquired through experience of dealing with differing runs of the same
code.

Application knowledge means one can sometimes achieve drastically superior
reduction in the total data compared with what one might achieve otherwise.
However, in the absence of such knowledge, we must equip the SSIO layers with
generic data reduction techniques and re-organization techniques.
Nevertheless, certain basic information on the semantics of the data is needed 
and must be supported by the overall infrastructure.

Another important consideration is knowledge regarding typical access patterns
for data in a given application. Scientific applications often read and analyze
data in a structured fashion and failure to respect this pattern, or to exploit
it through prefetching, can significantly increase the overall execution time
of the application. Accurate prefetching can only be achieved with access
pattern analysis. This requires an SSIO layer which can prefetch data in a
asynchronous fashion to take the bytes on the file.
